Let’s take a look at the file Wastes.csv. 
```{r eval = FALSE} 
# Change directory to file location, sourced from: https://stackoverflow.com/a/74626705
library(this.path)
setwd(this.path::here())

wastes <- read.csv("Wastes.csv")
```
1.  How many unique “Category” values are there in the data file (6 marks)? 
```{r eval = FALSE} 
unique_categories <- unique(wastes$Category)
print(length(unique_categories))
```
2.  How  many  negative  feedback  comments  are  in  the  'Description'  column  with  an 
environmental impact score of 2 or 3 (4 marks)? 
```{r eval = FALSE} 
library(dplyr)
negative_env_score <- length(grep(" 3/10.| 2/10.", wastes$Description))
print(negative_env_score)
```
3.  For each Category value, write code to calculate the fractions of waste tonnes of different 
waste sources, and then draw a chart to visualise the fraction numbers specific to the Category 
value (3 marks). 
```{r eval = FALSE} 

# Calculate the sum of 'value' for each category
category_sums <- aggregate(Tonnes ~ Category, data = wastes, sum)

# Print the category sums (optional, but good to check)
print(category_sums)
```
4.  Please add the 'Year' and 'State' values from Year_State_ID.csv to Wastes.csv, compute the 
total waste tonnes for each year and state, and store the result in a dataset named 'temp'. Then, 
use a single R function/command to display the statistical information (i.e., Min, Max, and 
Mean)  of  the  total  waste  tonnes  for  each  state  in  ‘temp’  (Note: You may use multiple 
functions/commands  to  prepare  the pre-processed data table, but when you compute and 
display the statistical information, you need to use a single R function/command.) (6 marks). 
```{r eval = FALSE} 

```
5.  Write code to draw a chart showing a yearly trend of total waste tonnes of food organics for 
each state. To draw the chart, please convert the Year-Year formats of all Year values into Year 
formats (3 marks).  
```{r eval = FALSE} 

```
6.  Write code to display the most recycled waste Type and the most disposed waste Type with the 
corresponding year. Write code to display the most increased waste Type over years (4 marks). 
```{r eval = FALSE} 

```
7.  Please investigate the factors influencing environmental impact scores. (Note: You may use 
variables  from  both  Wastes.csv  and  Year_State_ID.csv  for  your analysis.) Marks will be 
awarded based on the depth of your investigation (including analyses and discussions) and the 
strength of your reasoning (how insights and conclusions are drawn) (5 marks). 
```{r eval = FALSE} 

```
8.  Write code to only keep data records where the Category value is Hazardous wastes, the Type 
value is Tyres (T140) and  tonnes value is more than 0, then write code to add a new column 
named  “Tonnes_range”  and  fill  it  with  one  of  the  following  values  based  on  the 
corresponding “Tonnes” value: 
○        [0, 10000) 
○        [10000, 20000) 
○        [20000, 40000) 
○        [40000, 80000] 

Then, for each state, display a chart to show the number of cases of different score_range 
values. What do you observe? (4 marks) 
```{r eval = FALSE} 

```
9.  Using both original datasets and external sources, please investigate the factors influencing 
the yearly trend of total C&D waste tonnes. Marks will be awarded based on the depth of your 
investigation (including analyses and discussions) and the strength of your reasoning (how 
insights and conclusions are drawn). (Note: Please provide the source links for any external 
data used) (7 marks). 
```{r eval = FALSE} 

```